---
title: "TroyetalAnalysis1"
author: "KaelaReisdorf"
date: "2024-11-26"
output: html_document
---
#Data Cleaning & Set up for Analysis
```{r}
TroyetalFinalData <-read.csv("C:/Users/kaela/Downloads/TroyetalFinalData.csv", header=TRUE)
View(TroyetalFinalData)

```

```{r}
library(dplyr)
library(ggplot2)
library(emmeans)
```

##Creating my own subset of the data to view only what is needed for my analysis:
```{r}
colnames(TroyetalFinalData)

troyAnalysis<- subset(TroyetalFinalData [,c(1:29, 35:49)])

```


##Renamed columns:
```{r}
colnames(troyAnalysis) <- c("participant ID", "dateOfTest", "dobMonth","dobDay", "dobYear","raceEthnicity", "raceEthnicityDescription", "gender", "genderDescription", "income", "CRA1", "CRA2", "CRA3", "CRA4", "CRA5", "CRA6", "CRA7", "CRA8",
"HCRA1", "HCRA2", "HCRA3", "HCRA4", "HCRA5", "HCRA6", 
"DS1", "DS2", "DS3", "DS4", "DS5",
"M1","M2","M3","M4","M5","M6", "M7", "M8", "M9", "M10", "M11", "M12", "M13", "M14", "M15")
```

##Change particpant answers to numerical values:
```{r}
troyAnalysis <- troyAnalysis %>%
  mutate(across(c("CRA1", "CRA2", "CRA3", "CRA4", "CRA5", "CRA6", "CRA7", "CRA8","HCRA1","HCRA2","HCRA3","HCRA4","HCRA5","HCRA6"), as.character))

##Cognitive Reappraisal, Habitual Cognitive Reappraisal, and Depressive Symptoms##
troyAnalysis <- troyAnalysis %>%
  mutate(across(c("CRA1","CRA2","CRA3","CRA4","CRA5","CRA6","CRA7","CRA8"), ~ case_when(
    . == "Strongly disagree" ~ 1,
    . == "Disagree" ~ 2,
    . == "Somewhat disagree" ~ 3,
    . == "Neutral" ~ 4,
    . == "Agree" ~ 5,
    . == "Somewhat agree" ~ 6,
    . == "Strongly agree" ~ 7,
    TRUE ~ NA_real_
  )), across(c("HCRA1","HCRA2","HCRA3","HCRA4","HCRA5","HCRA6"),
                ~ case_when(
                  . == "Strongly disagree" ~ 1,
                  . == "Disagree" ~ 2,
                  . == "Somewhat disagree" ~ 3,
                  . == "Neither agree nor disagree" ~ 4,
                  . == "Agree" ~ 5,
                  . == "Somewhat agree" ~ 6,
                  . == "Strongly agree" ~ 7,
                  TRUE ~ NA_real_ 
  )), across(c("DS1", "DS2", "DS3", "DS4", "DS5"),
                ~ case_when(
                  . == "Rarely or none of the time (less than 1 day)" ~ 1,
                  . == "Some or a little of the time (1-2 days)" ~ 2,
                  . == "Occasionally or moderate amount of time (3-4 days)" ~ 3,
                  . == "Most or all of the time (5-7 days)" ~ 4,
                  TRUE ~ NA_real_)))

```

```{r}
##Mindfulness##
troyAnalysis <- troyAnalysis %>%
  mutate(across(c("M1","M2","M3","M4","M5","M6", "M7", "M8", "M9", "M10", "M11", "M12", "M13", "M14", "M15"),
                ~ case_when(
                  . == "Almost Never" ~ 1,
                  . == "Very Infrequently" ~ 2,
                  . == "Somewhat Infrequently" ~ 3,
                  . == "Somewhat Frequently" ~ 4,
                  . == "Very Frequently" ~ 5,
                  . == "Almost Always" ~ 6, 
                  TRUE ~ NA_real_ )))
```



#Statistical Analysis Work


##Average scores for CRA, HCRA, DS, & M
```{r}
###If trouble with code, restart R - possible errors occur when using mutate + starts with functions
troyAnalysis <- troyAnalysis %>%
  mutate(
    crScore = rowMeans(select(., starts_with("CRA")), na.rm = TRUE),
    hcrScore = rowMeans(select(., starts_with("HCRA")), na.rm = TRUE),
    dsScore = rowMeans(select(., starts_with("DS")), na.rm = TRUE),
    mScore = rowMeans(select(., starts_with("M")), na.rm = TRUE),
    crCentered = crScore - mean(crScore, na.rm = TRUE),
    hcrCentered = hcrScore - mean(hcrScore, na.rm = TRUE),
    dsCentered = dsScore - mean(dsScore, na.rm = TRUE),
    mCentered = mScore - mean(mScore, na.rm = TRUE)
  )

##code with centered mean scores
```


##Research Question 1: How does mindfulness as a moderator influence the relationship between cognitive reappraisal and depression symptoms in individuals?
  ###Hypothesis: Individuals will show a stronger negative relationship between cognitive reappraisal and depressive symptoms as there levels of mindfulness increase.
```{r}
library(rempsyc)
library(easystats)
```

```{r}
troyModerations <- nice_mod(
  data = troyAnalysis,
  response = "dsCentered",
  predictor = "crCentered",
  moderator = "mCentered"
)
troyModerations
(modTable <- nice_table(troyModerations, highlight = TRUE))

##Significance between depressive symptoms & mindfulness
```

```{r}
troySlopes <- nice_slopes(
  data = troyAnalysis,
  response = "dsCentered",
  predictor = "crCentered",
  moderator = "mCentered"
)
troySlopes
nice_table(troySlopes, highlight = TRUE)
```

###Visual for research question 1:
```{r}
#defining model for plot function
moderateTroy1 <- lm(dsCentered ~ crCentered * mCentered, data = troyAnalysis)

#plotting the model
library(interactions)
interact_plot(moderateTroy1, pred = "crCentered", modx = "mCentered", interval = TRUE)
```


##Research Question 2: How do levels of HCRA mediate the association between mindfulness and depressive symptoms?
  ###Hypothesis: As HCRA and Mindfulness increase, depressive symptoms decrease.
  
```{r}
library(mediation)
#defining model HCRCA + Mindfulness -> Depressive Symptoms
troyMediator<-lm(hcrCentered ~ mCentered, data = troyAnalysis)

troyMediatorOutcome<-lm(dsCentered ~ mCentered + hcrCentered, data = troyAnalysis)
troyMediator

researchResult2 <- mediate(troyMediator, troyMediatorOutcome, treat = "mCentered", mediator = "hcrCentered", boot = TRUE)

summary(researchResult2)
```

###Visual for research question 2:
```{r}
ggplot(troyAnalysis, aes(x = mScore, y = dsScore)) +
  geom_point() +  
  # Direct path: Mindfulness -> Depressive Symptoms
  geom_smooth(method = "lm", color = "purple", se = TRUE, linetype = "dashed") +  
  # Indirect path: Mindfulness -> HCRA -> Depressive Symptoms
  geom_smooth(aes(x = mScore, y = hcrScore), method = "lm", color = "blue", se = FALSE, linetype = "solid") +  
  geom_smooth(aes(x = hcrScore, y = dsScore), method = "lm", color = "blue", se = FALSE, linetype = "solid") +
  labs(
    title = "HCRA as a Mediator for Mindfulness and Depressive Symptoms",
    x = "Mindfulness",
    y = "Depressive Symptoms",
    subtitle = "Purple dashed line = Direct Effect, Blue solid lines = Indirect Effect (Mediation)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```


##Research Question 3: To what extent do mindfulness and CRA independentaly predict depressive symptoms?
  ###Hypothesis: Mindfulness and CRA independently will decrease depressive symptoms.
  
```{r}
troyLM <- lm(dsCentered ~ mCentered + crCentered, data = troyAnalysis)

summary(troyLM)
```

###Visual Mindfulness->Depressive Symptoms
```{r}
ggplot(troyAnalysis, aes(x = mCentered, y = dsCentered)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "purple", linetype = "solid") +
  labs(title = "Relationship Between Mindfulness and Depressive Symptoms",
       x = "Mindfulness",
       y = "Depressive Symptoms") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

###Visual CRA -> Depressive Symptoms
```{r}
ggplot(troyAnalysis, aes(x = crCentered, y = dsCentered)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "purple", linetype = "solid") +
  labs(title = "Relationship Between Cognitive Reappraisal and Depressive Symptoms",
       x = "Cognitive Reappraisal",
       y = "Depressive Symptoms") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```


```{r}
write.csv(troyAnalysis, "troyAnalysis_subset.csv", row.names = FALSE)
```

